---
title: "'ggplot' 1"
author: "Katie"
date: "10/12/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) 
library(ggridges) 
```

## Create the weather data 

```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Create a ggplot 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point() 
```

## Alternate way of making this plot 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point() 
```

## Saving initial plots (Jeff mostly doesn't do it this way) 

```{r}
scatterplot = 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point() 

scatterplot 
```

## Adding color 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4)  
```

## Why do 'aes' positions matter? 

Here's the first plot: 
```{r}
## this makes one smooth line without the gray uncertainty band around it
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4) + 
  geom_smooth(se = FALSE)  
```

vs. 

The second plot: 
```{r}
## this makes 3 different colored smooth lines for each of the locations 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .4) + 
  geom_smooth(se = FALSE) 
```

## Time to facet! 

```{r}
## creates 3 smooth lines that are the same color as the locations so is kinda hard to see
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .4) + 
  geom_smooth(se = FALSE) + 
  facet_grid(~name)

## creates 3 smooths lines that stand out better amongst the locations and is easier to see  
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4) + 
  geom_smooth(se = FALSE) + 
  facet_grid(~name)
```

## This is fine. But not interesting. 

```{r}
weather_df %>% 
  ggplot(aes(x= date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .35) + 
  geom_smooth(size = 2, se = FALSE)
```

Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()). 

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = (tmax * 9 / 5) + 32, 
    tmin= (tmin * 9 / 5) + 32) %>% 
ggplot(aes(x = tmin ,y = tmax)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = FALSE) 

## Answer Key Solution  
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(tmax_fahr = tmax * (9 / 5) + 32,
         tmin_fahr = tmin * (9 / 5) + 32) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)
```

## Some extra stuff 

```{r}

```





